A Test of the Frequency Independence Assumption of the Power System Parameters used in Geomagnetically Induced Current Calculations

### 2019-02-01 paper.tex
Updates from ShareLaTeX

One-second-cadence geoelectric field and geomagnetic field data from Kakioka Magnetic Observatory were obtained from the Japan Meteorological Agency data portal. Data for the longest continuous time span, 2006-04-02 through 2006-04-10 are shown in the top and middle panels of Figure~\ref{figure1}.

Unphysical spikes in the electric field measurements assumed to be unphysical when the absolute value of a component $i$ of the electric field changed by more than 1 mV/km at time $t_s$, and the values $E_i(t_s-2)$ through $E_i(t_s+4)$ were replaced using linear interpolation of the values at $E_i(t_s-3)$ and $E_i(t_s + 5)$. In the sample shown in Figure~\ref{figure1}, 21 spikes were identified in $E_x$ and zero in $E_y$.

@@

The GIC dataset includes 1-second raw data and 1 Hz low-pass filtered raw measurements.  The results presented here are for the 1 Hz low-pass filtered GIC values and data for the time interval 2006-04-02 through 2006-04-10 is shown in the bottom panel of Figure~\ref{figure1}.

The GIC data contains non-physical spikes that are followed by a relaxation time of approximately 80 seconds. These spikes were removed by identifying times $t_s$ when the change in $|GIC|$ was greater than $0.04$ Amps and then data in a window $GIC(t_s-2)$ through $GIC(t_s+100)$ was replaced using a linear interpolation of the values at $GIC(t_s-3)$ and $GIC(t_s + 101)$. In the example interval shown in Figure~\ref{figure1}, 493 spikes were removed.

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{figures/plot_raw_All_20060402.pdf}
\caption{(Top) One-second-cadence geomagnetic field measured at the Memanbetsu Magnetic Observatory (MMB). (Middle) One-second-cadence geoelectric field measured at the Memanbetsu Magnetic Observatory (MMB) $\sim$10 km from the Memanbetsu substation. (Bottom) One-second-cadence GIC measurements measured in a grounded neutral point of a Y-connected three-phase transformer connected to a 187 kV bus at the Memanbetsu substation of the Hokkaido Power Co. The subscripts $x$, $y$, and $z$ correspond to North, East, and Downward, respectively.}
\label{figure1}
\end{figure}

@@

The average MSE ratio was computed by averaging the 35 MSE ratios and the average signal-to-noise ratio at a given frequency is the average of the 35 signal-to-noise ratios at that frequency.

The 95\% confidence interval, CI, on the  metrics was calculated using 1000 bootstrap samples of the 35 prediction metrics. The typical difference between the limits when the confidence intervals were computed assuming the values were gaussian-distributed was $\sim$2%.

\begin{table}
\caption{Model performance metrics for models $m=1-4$.}
\centering
\begin{tabular}{l c c c c}
\hline
$m$\hspace{1em} Model & PE & 95\% CI & MSE$_m$/MSE$_1$ & 95\% CI\\
\hline
1\hspace{1em} $G_o(t) = a_oE_x(t) + b_oE_y(t)$ & 0.37 & [0.31, 0.43] & 1.0 & \\
2\hspace{1em} $G_E(t) = A(\omega)E_x(t) + B(\omega)E_y(t)$ & 0.60 & [0.54, 0.65] & 1.8 & [1.6,2.0]\\
3\hspace{1em} $G_{E'}(t) = A(\omega)E'_x(t) + B(\omega)E'_y(t)$ & 0.78 & [0.76, 0.80] & 3.1 & [2.8,3.4]\\
4\hspace{1em} $G_{B}(t) = Z_xB_x(t) + Z_yB_y(t)$ & 0.83 & [0.80, 0.85] & 4.6 & [3.8,5.3]\\
\hline
\end{tabular}
\end{table}

The primary conclusion from the results shown in Table~1 is that the frequency dependent model, Model 2, is significantly better at predicting $GIC$ than the frequency independent model, Model 1.
Model 3, which uses magnetic field measurements as in input provides better estimates of $GIC$ than Model 2. The last conclusion may be somewhat unexpected because Model 3 uses parameters from two different models. A proposed explanation is that the electric field measurements contain more noise than the magnetic field measurements.

The average parameters for Model 1 are $a_o =$ $71.9 \pm 27.8$ A/(V/km) and $b_o =$ $-70.0 \pm 65.8$ A/(V/km) and the distribution of the 35 parameters calculated using each 1-day interval is shown in Figure~2. The wide distribution and large error bars are suggestive that the model is a poor representation of the measurements. However, the model does have a good prediction performance - the PE is 0.37 and the data/model correlation coefficient of 0.64. In a similar way that $|\mathbf{B}(t)|$ or $d|\mathbf{B}(t)|/dt$ are often good predictors of $GIC(t)$ [REFS] despite the fact that physically $\mathbf{E}$ drives $GIC$ and $\mathbf{E}$ is related to $\mathbf{B}$ through a frequency dependent transfer function, Model 1 is a fair predictor of $GIC$, but is a poor approximation relative to Models 2 and 3.

The average parameters for Models 2 and 4 are shown in Figure~\ref{H}--\ref{Phi}. The error bars at each frequency or time is the standard error of the 35 values used to compute the average. Figure \ref{H} shows that the timescale of variations to impulse in $\mathbf{E}$ for Model 2 and $\mathbf{B}$ for Model 4 are similar.

Figure~\ref{Z} shows that the frequency domain transfer functions for Model 2 vary by an order of magnitude on timescales of 60 seconds to 12 hours (that for Model 1 is constant by definition).

@@

\item Averaging the 35 transfer functions using the Huber M-estimator with a psi-function \citep{Huber2011}.

@@

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{figures/plot_TF_aves_SN-options-1.pdf}
\caption{Signal-to-Noise}
\label{SN}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{figures/plot_TF_aves_aobo_histograms-options-1.pdf}
\caption{Distribution of parameters in Model 1. Each set of parameters were computed 35 times using one day of 1-second cadence data. The dots indicate the average of each histogram: $a_o$ = 71.9 A/(V/km) and $b_o$ = -70.0 A/(V/km).}
\label{histogram}
\end{figure}

@@

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{figures/plot_TF_aves_Phi-options-1.pdf}
\caption{Frequency domain phase values for Models 1-3. The phase at each frequency for parameter $p$ is arg$(p)$.}
\label{Phi}
\end{figure}

>>>>

One-second-cadence geoelectric field and geomagnetic field data from Kakioka Magnetic Observatory were obtained from the Japan Meteorological Agency data portal. Data for the longest continuous time span, 2006-04-02 through 2006-04-10 are shown in the top and middle panels of Figure~\ref{sample}.

Unphysical spikes in the electric field measurements assumed to be unphysical when the absolute value of a component $i$ of the electric field changed by more than 1 mV/km at time $t_s$, and the values $E_i(t_s-2)$ through $E_i(t_s+4)$ were replaced using linear interpolation of the values at $E_i(t_s-3)$ and $E_i(t_s + 5)$. In the sample shown in Figure~\ref{sample}, 21 spikes were identified in $E_x$ and zero in $E_y$.

@@

The GIC dataset includes 1-second raw data and 1 Hz low-pass filtered raw measurements.  The results presented here are for the 1 Hz low-pass filtered GIC values and data for the time interval 2006-04-02 through 2006-04-10 is shown in the bottom panel of Figure~\ref{sample}.

The GIC data contains non-physical spikes that are followed by a relaxation time of approximately 80 seconds. These spikes were removed by identifying times $t_s$ when the change in $|GIC|$ was greater than $0.04$ Amps and then data in a window $GIC(t_s-2)$ through $GIC(t_s+100)$ was replaced using a linear interpolation of the values at $GIC(t_s-3)$ and $GIC(t_s + 101)$. In the example interval shown in Figure~\ref{sample}, 493 spikes were removed.

@@

The average PE and MSE ratio was computed by averaging the 35 PE and MSE ratios; the average signal-to-noise ratio at a given frequency is the average of the 35 signal-to-noise ratios at that frequency. The 95\% confidence interval, CI, on the PE and MSE ratio was calculated using 1000 bootstrap samples of the 35 prediction metrics. The typical difference between the limits when the confidence intervals were computed assuming the values were Gaussian-distributed was $\sim$2\%.

\begin{table}
\caption{Out-of-sample model performance metrics for models $m=1-4$. $MSE_m/MSE_1$ is the ratio of the MSE for model $m$ to that of Model 1. The PE and MSE ratios are the average of 35 values and the confidence interval, CI, was determined from 1000 bootstrap samples of these 35 values.}
\centering
\begin{tabular}{l c c c c}
\hline
$m$\hspace{1em} Model & PE & 95\% CI & MSE$_m$/MSE$_1$ & 95\% CI\\
\hline
1\hspace{1em} $G_o(t) = a_oE_x(t) + b_oE_y(t)$ & 0.35 & [0.25, 0.45] & 1.0 & \\
2\hspace{1em} $G_E(t) = A(\omega)E_x(t) + B(\omega)E_y(t)$ & 0.60 & [0.54, 0.65] & 1.8 & [1.6,2.0]\\
3\hspace{1em} $G_{E'}(t) = A(\omega)E'_x(t) + B(\omega)E'_y(t)$ & 0.78 & [0.76, 0.80] & 3.1 & [2.8,3.4]\\
4\hspace{1em} $G_{B}(t) = Z_xB_x(t) + Z_yB_y(t)$ & 0.83 & [0.80, 0.85] & 4.6 & [3.8,5.3]\\
\hline
\end{tabular}
\end{table}

Figure~\ref{predictions} shows a 1-day interval for which the value of the prediction efficiency for each model was in the same order as that shown in Table~1, i.e., Model 4 has the highest PE and Model 1 has the lowest.

The primary conclusions of this work follow from the results shown in Table~1: the frequency dependent model, Model 2, is significantly better at predicting $GIC$ than the frequency independent model, Model 1. Model 3, which uses magnetic field measurements as in input provides better estimates of $GIC$ than Model 2. Model 4, in which a transfer function that connects $GIC$ to $\mathbf{B}$ was derived directly from the data, provides the best predictions of $GIC$.

Figure~\ref{SN} shows the average signal-to-noise ratios as a function of period for the four models. The error bars represent one standard error of the 35 SN ratios at each period. Consistent with the results in Table~1, the SN ratio for Model 1 is consistently lower than that for Models 2-4.  Above $1400$~s (24~min), the SN ratio ordering is consistent with the PE and MSE ratios shown in Table~1. Below $500$~s (8~min), Model 2 has the largest SN ratio. This result is somewhat visible in the sample interval shown in Figure~\ref{predictions} - at approximately 2006-08-07 07:00, the high frequency fluctuations in the prediction error for Model 2 are smaller than that for Models 3 and 4.

The average parameters for Model 1 are $a_o =$ $71.9 \pm 27.8$ A/(V/km) and $b_o =$ $-70.0 \pm 65.8$ A/(V/km) and the distribution of the 35 parameters calculated using each 1-day interval is shown in Figure~2. The wide distribution and large error bars are suggestive that the model is a poor representation of the measurements. However, the model does have a good prediction performance - the PE is 0.35 and the data/model correlation coefficient of 0.64. In a similar way that $|\mathbf{B}(t)|$ or $d|\mathbf{B}(t)|/dt$ are often good predictors of $GIC(t)$ [REFS] despite the fact that physically $\mathbf{E}$ drives $GIC$ and $\mathbf{E}$ is related to $\mathbf{B}$ through a frequency dependent transfer function, Model 1 is a fair predictor of $GIC$, but is a poor approximation relative to Models 2 and 3. For the data considered in this work, when a model $G(t) = c_oB_x(t) + d_oB_y(t)$ is used, were $c_o$ and $d_o$ are constants, the in-sample average prediction efficiency is comparable to that of Model 1, but the out-of-sample prediction efficiency is negative.

The average parameters for Models 2 and 4 are shown in Figure~\ref{H}--\ref{Phi}. The error bars at each time step or period is the standard error of the 35 values used to compute the average.

Figure \ref{H} shows that the timescale of variations of an impulse in $\mathbf{E}$ for Model 2 and $\mathbf{B}$ for Model 4 are similar and are different from zero at more than one time value (the impulse response for Model 3 is similar to that of Model 2 and is not shown). By definition, the impulse response for Model 1 is zero except at $t=0$. Note that the sign of the impulse response at $t=0$ is opposite for Model 1 and Model 2 for the $E_y$ coefficient - at $t=0$, $b_o < 0$ while $b(0)>0$.

Figure~\ref{Z} shows that the frequency domain transfer functions for Model 2 vary by an order of magnitude on timescales of 60 seconds to 12 hours (that for Model 1 is constant by definition). For this system, GIC is most dependent on $E_y$, which has less variability than $E_x$ - the variance in $E_x$ is $\sim$5x larger than $E_y$ over the entire dataset (this is also visible in Figure~\ref{sample}).

@@

\item Averaging the 35 transfer functions using the Huber M-estimator with a $\psi$-function \citep{Huber2011}.

@@

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{figures/plot_raw_All_20060402.pdf}
\caption{(Top) One-second-cadence geomagnetic field measured at the Memanbetsu Magnetic Observatory (MMB). (Middle) One-second-cadence geoelectric field measured at the Memanbetsu Magnetic Observatory (MMB) $\sim$10 km from the Memanbetsu substation. (Bottom) One-second-cadence GIC measurements measured in a grounded neutral point of a Y-connected three-phase transformer connected to a 187 kV bus at the Memanbetsu substation of the Hokkaido Power Co. The subscripts $x$, $y$, and $z$ correspond to North, East, and Downward, respectively.}
\label{sample}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{figures/plot_model_predictions-MeanModel-20060806T150000.pdf}
\caption{Out-of-sample model predictions and predictions errors for a 1-day interval. The bottom overlapping traces are the measured GIC (black line) and the model predictions (colored lines). The prediction error of each model is shown shifted by an arbitrary offset.}
\label{predictions}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{figures/plot_TF_aves_SN-options-1.pdf}
\caption{Signal-to-noise ratios at the evaluation frequencies for the four models.}
\label{SN}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{figures/plot_TF_aves_aobo_histograms-options-1.pdf}
\caption{Distribution of parameters in Model 1. Each set of parameters were computed 35 times using one day of 1-second cadence data. The circle/dot on the horizontal axis indicates the average of each histogram: $a_o$ = 71.9 A/(V/km) and $b_o$ = -70.0 A/(V/km).}
\label{histogram}
\end{figure}

@@

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{figures/plot_TF_aves_Phi-options-1.pdf}
\caption{Frequency domain phase values for Models 1-3.}
\label{Phi}
\end{figure}
